Cores from kicks: inferring the gravitational recoil history of SMBHs in elliptical galaxies.

### 2024-05-25 papers/paper-corekick/paper-corekick.tex
New sketch figures for Atte

\subsection{Merger Simulations}\label{ssec:mergers}
We merge two independently Monte Carlo sampled galaxy ICs, creating an equal mass galaxy merger with a near-radial orbit. 
The merger orbit has an initial separation of $D=30\,\mathrm{kpc}$, a first pericentre distance of $r_\mathrm{peri}=2\,\mathrm{kpc}$, and an initial eccentricity of $e_0=0.97$.
The merger configuration results in a rapid coalescence of the BH binary, however we note that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
We then select the snapshot just prior ($\sim8\,\mathrm{Myr}$) to the GW-driven SMBH merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors. To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
At the moment of the merger, the centre of mass (CoM) velocity of the SMBH binary is $v_{\mathrm{bin},\mathrm{CoM}}\sim[4.2, 1.1, 1.6]\,\kmps$ in the global coordinate frame.
The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$.
As discussed in \drafting{section}, higher recoil velocities typically result in larger excursions of the remnant SMBH from the isophotal centre.
To provide a fair comparison between simulations, we choose to analyse each child simulation at a fixed time after the recoil kick, namely $t\simeq 0.075\,\mathrm{Gyr}$.
This value is chosen so as to minimise the effects of relaxation in the central regions of the galaxy remnant.
We estimate the relaxation time $t_\mathrm{relax}$ at a radius $r$ as \citep{binney2008}:
\begin{equation}\label{eq:relax}
    t_\mathrm{relax} \simeq 2.1 \frac{\sigma r^2}{G \bar{m} \ln \Lambda},
\end{equation}
where $\sigma$ is the particle velocity dispersion within $r$, $\bar{m}$ is the mean particle mass (hence the mean of the stellar and DM particle masses), and $\Lambda$ is the argument of the Coulomb logarithm, given by:
\begin{equation}
    \Lambda = \frac{r \langle v^2 \rangle}{2 G \bar{m}},
\end{equation}
where $\langle v^2 \rangle$ is the mean squared particle velocity within $r$. 
The relaxation time increases quadratically with increasing $r$, and we find for a radius of $r=0.1\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 0.1 \, \mathrm{Gyr}$, thus motivating our choice of $0.075\,\mathrm{Gyr}$ for the analysis time. 
We  note that \autoref{eq:relax} does not take into account gravitational softening, and thus provides a conservative lower bound for the relaxation time.
This is critical as the \ketju{} region does not include gravitational softening between stellar particles and the SMBH, so is a not fully-softened simulation.

@@

\subsection{Infall time}
\drafting{
    \textbf{Atte's section}. Would be good to have:
    \begin{enumerate}
        \item Plot of infall time (i.e., time for stability criterion to be reached) as a function of kick velocity
        \item Number of oscillations before settling?
        \item Plot of maximum displacement of kicked BH as a function of kick velocity. How does this compare to core size? Which simulations have BH kicked beyond the core?
    \end{enumerate}
}
\drafting{
    Include this text here, this is how we identify `settled' systems, but we're not really interested in when the SMBH has settled for the rest of the paper (refer to Nasim where the majority of the GW-induced scouring occurs during the first excursion of the SMBH).
    Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
    The time to this point, from the time of SMBH binary coalescence, we term the `infall time'.
    Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
}

\drafting{
    Also maybe we can discuss the number of bound particles to the BH here. I've already looked at this, but we can talk about combining that information with the apocentre plots. 
    It seems that increasing $\vk$ decreases the amount of bound mass, in agreement with previous studies, but also for each $\vk$, the maximum amount of stellar mass bound to the BH is maximal at apocentre.
}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{infall_time}
    \caption{Plot of infall time (time to stability criterion) as a function of kick velocity.}
    \label{fig:infall}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{max_disp}
    \caption{Plot of the maximum BH displacement as a function of kick velocity. Where do the displacement (red) and core size (green) lines intersect?}
    \label{fig:max_disp}
\end{figure}

>>>>

\subsection{Merger Simulations}\label{ssec:mergers}
We merge two independently Monte Carlo sampled galaxy ICs, creating an equal mass galaxy merger with a near-radial orbit. 
The merger orbit has an initial separation of $D=30\,\mathrm{kpc}$, a first pericentre distance of $r_\mathrm{peri}=2\,\mathrm{kpc}$, and an initial eccentricity of $e_0=0.97$.
The merger configuration results in a rapid coalescence of the BH binary, however we note that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
We then select the snapshot just prior ($\sim8\,\mathrm{Myr}$) to the GW-driven SMBH merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors. To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
At the moment of the merger, the centre of mass (CoM) velocity of the SMBH binary is $v_{\mathrm{bin},\mathrm{CoM}}\sim[4.2, 1.1, 1.6]\,\kmps$ in the global coordinate frame.
The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$.
As discussed in \drafting{section}, higher recoil velocities typically result in larger excursions of the remnant SMBH from the isophotal centre.
\drafting{Need to edit all this, \textbf{Atte} could you do this seeing as you've chosen the snapshots?

To provide a fair comparison between simulations, we choose to analyse each child simulation at a fixed time after the recoil kick, namely $t\simeq 0.075\,\mathrm{Gyr}$.
This value is chosen so as to minimise the effects of relaxation in the central regions of the galaxy remnant.
We estimate the relaxation time $t_\mathrm{relax}$ at a radius $r$ as \citep{binney2008}:
\begin{equation}\label{eq:relax}
    t_\mathrm{relax} \simeq 2.1 \frac{\sigma r^2}{G \bar{m} \ln \Lambda},
\end{equation}
where $\sigma$ is the particle velocity dispersion within $r$, $\bar{m}$ is the mean particle mass (hence the mean of the stellar and DM particle masses), and $\Lambda$ is the argument of the Coulomb logarithm, given by:
\begin{equation}
    \Lambda = \frac{r \langle v^2 \rangle}{2 G \bar{m}},
\end{equation}
where $\langle v^2 \rangle$ is the mean squared particle velocity within $r$. 
The relaxation time increases quadratically with increasing $r$, and we find for a radius of $r=0.1\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 0.1 \, \mathrm{Gyr}$, thus motivating our choice of $0.075\,\mathrm{Gyr}$ for the analysis time. 
We  note that \autoref{eq:relax} does not take into account gravitational softening, and thus provides a conservative lower bound for the relaxation time.
This is critical as the \ketju{} region does not include gravitational softening between stellar particles and the SMBH, so is a not fully-softened simulation.
}

@@

\subsection{Infall time}
\drafting{
    \textbf{Atte's section}. Would be good to have:
    \begin{enumerate}
        \item Plot of infall time (i.e., time for stability criterion to be reached) as a function of kick velocity
        \item Number of oscillations before settling, probably on the same plot but as a second y axis?
        \item Plot of apocentre of kicked BH as a function of kick velocity. How does this compare to core size? Which simulations have BH kicked beyond the core? Can fit a function to data maybe
        \item Langrangian radii: how is the stellar mass moving about? 
    \end{enumerate}
}
\drafting{
    Include this text here, this is how we identify `settled' systems, but we're not really interested in when the SMBH has settled for the rest of the paper (refer to Nasim where the majority of the GW-induced scouring occurs during the first excursion of the SMBH).
    Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
    The time to this point, from the time of SMBH binary coalescence, we term the `infall time'.
    Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
}

\drafting{
    Also maybe we can discuss the number of bound particles to the BH here. I've already looked at this, but we can talk about combining that information with the apocentre plots. 
    It seems that increasing $\vk$ decreases the amount of bound mass, in agreement with previous studies, but also for each $\vk$, the maximum amount of stellar mass bound to the BH is maximal at apocentre.
}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{infall_apo}
    \caption{Plot of infall time (time to stability criterion) as a function of kick velocity, and number of oscillations. Maybe can do apocentre plot too: make this a 2x1 figure??}
    \label{fig:infall}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{lagrangian}
    \caption{Plot of Lagragian radii: comment on matter movement}
    \label{fig:max_disp}
\end{figure}
